Supermassive black holes and their host spheroids \\ III. A universal $M_{\rm BH} - n_{\rm sph}$ correlation

### 2015-12-02 papers/mbhn/mbhn.tex
g

\section{Data}
We populated the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams 
with the same galaxy sample used in \emph{Paper II} (and presented here in Table \ref{tab:sample}), 
i.e.~66 galaxies for which a dynamical measurement of the black hole mass has been reported in the literature 
(by \citealt{grahamscott2013} or \citealt{rusli2013}) 
and for which we were able to successfully model the light distribution and measure the spheroid structural parameters 
using $3.6\rm~\mu m$ \emph{Spitzer} satellite images. 
Our galaxy decompositions take into account bulge, disks, spiral arms, bars, rings, halo, 
extended or unresolved nuclear source and partially depleted core, 
and -- for the first time -- they were checked to be consistent with the galaxy kinematics 
\citep{atlas3dIII,scott2014,arnold2014}. 
Kinematical information was used to confirm the presence of disk components 
in the majority of early-type (elliptical + lenticular) galaxies and, more importantly,  
to estabilish the radial extent of these disks, 
which in most cases is not obvious from a visual inspection of the galaxy images. 
This enabled us to distinguish between intermediate-scale disks, 
that are fully embedded in the spheroid,  
and large-scale disks, that encase the bulge and dominate the light at large radii.  
\cite{ellicular} demonstrate that when an intermediate-scale disk is misclassified and modeled as a large-scale disk, 
the luminosity of the spheroid is underestimated, 
hence the galaxy incorrectly appears as a positive outlier (an ``over-massive'' black hole) in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
A detailed description of the dataset used here, the data reduction process and the galaxy modelling technique that we developed 
can be found in \emph{Paper I}, 
along with a discussion of how we estimated the uncertainties on the spheroid S\'ersic indices\footnote{The uncertainties associated with 
the spheroid S\'ersic indices were estimated with a method that takes into account systematic errors. 
This method consists in comparing, for each of our galaxies, the measurements of the spheroid S\'ersic index obtained by other studies 
with that obtained by us. 
Systematic errors are typically not considered by popular 2D fitting codes, which report only the statistical errors 
associated with their fitted parameters. 
Readers should refer to \emph{Paper I} for a more detailed discussion on this topic. }.  
The morphological classification (E = elliptical; E/S0 = elliptical/lenticular; S0 = lenticular; 
S0/Sp = lenticular/spiral; Sp = spiral; and ``merger'') follows from the galaxy decompositions illustrated in \emph{Paper I}. 
As in \emph{Paper II}, we will refer to early-type galaxies (E+S0) and late-type galaxies (Sp). 
The early-type bin includes the two galaxies classified as E/S0, 
whereas the two galaxies classified as S0/Sp and the two galaxies classified as ``mergers'' are included in neither the early- nor the late-type bin.

@@

\section{Analysis and results}
\label{sec:anal}
As in \emph{Paper II}, a linear regression analysis of the $L_{\rm sph} - n_{\rm sph}$ 
(Table \ref{tab:lregLn} and Figure \ref{fig:magn})
and $M_{\rm BH} - n_{\rm sph}$ (Table \ref{tab:lregMn} and Figure \ref{fig:mbhn}) diagrams 
was performed using three different routines: 
the BCES code from \cite{akritasbershady1996}, 
the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}.
All of these three routines take into account the intrinsic scatter, 
but only the FITEXY and the {\tt linmix\_err} codes allow one to quantify it.
\cite{tremaine2002} cautioned that 
the BCES estimator becomes ineffective when the dataset contains at least one low-precision measurement -- 
regardless of how many high-precision measurements are in the sample -- 
and tends to be biased in case of low number statistics 
or if the mean square of the uncertainties associated to the independent variable is comparable to 
the variance of the distribution of the independent variable. 
According to the results from the Monte Carlo Markov Chain simulations of \cite{tremaine2002} and \cite{novak2006}, 
these problems can be overcome with the use of the modified FITEXY routine. 
\cite{park2012} also concluded that the modified FITEXY routine performs better and returns less biased results than the BCES estimator, 
and noted that the the modified FITEXY routine is computationally less intensive than the Bayesian technique {\tt linmix\_err}. 
Given that at least one of our subsamples (the lenticular galaxies) has a small size 
and that the uncertainties associated to $n_{\rm sph}$ are relatively large compared to the range in $n_{\rm sph}$ values for most of our subsamples, 
we put more trust in the results obtained with the modified FITEXY routine 
and throughout the text we quote only those. \\ 
We report both symmetrical and nonsymmetrical linear regressions.  
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas nonsymmetrical forward ($Y|X$) regressions -- which minimize the scatter in the $Y$ direction -- 
allow one to predict the value of the observable $Y$ with the best possible precision. \\
We searched for extreme outliers in both the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams, 
and found that in our $L_{\rm sph} - n_{\rm sph}$ plot there are no $3\sigma$ outliers, 
whereas in our $M_{\rm BH} - n_{\rm sph}$ plot 
the lenticular galaxies NGC 0524 and NGC 3998 reside more than $3\sigma$ from the bisector linear regression for all galaxies. 
These two galaxies have therefore been excluded from the rest of the analysis. \\

\subsection{$L_{\rm sph} - n_{\rm sph}$}
Following \cite{graham2001bulges}, 
who showed that the $L_{\rm sph} - n_{\rm sph}$ relation is different for elliptical galaxies and the bulges of disk galaxies (S0+Sp), 
\cite{savorgnan2013} re-analyzed the data from \cite{grahamguzman2003} and \cite{grahamworley2008} 
and obtained two separate $L_{\rm sph} - n_{\rm sph}$ linear regressions for elliptical galaxies and the bulges of disk galaxies 
(in the B- and K-band, respectively). 
At the time, the $L_{\rm sph} - n_{\rm sph}$ datasets from \cite{grahamguzman2003} and \cite{grahamworley2008} were of the best quality available 
to investigate the $L_{\rm sph} - n_{\rm sph}$ relation for different galaxy morphological types. 
However, these datasets were not obtained from a homogeneous analysis, 
but they were a collection of results taken from various past bulge/disk decomposition studies. 
%It is also worth mentioning that, 
%although the K-band luminosities of the bulges of disk galaxies had been corrected for internal dust absorption, 
%{\bf such corrections are still not very accurate today?!? Driver et al. xxxx??? }  
Here we re-investigate the $L_{\rm sph} - n_{\rm sph}$ diagram using only our high-quality dataset. 
Spheroid luminositites and S\'ersic indices were obtained from accurate multicomponent decompositions,  
performed in a consistent manner using the $3.6\rm~\mu m$ band, which is less affected by dust extinction than the K-band. \\
\cite{grahamworley2008} presented a single $L_{\rm sph} - n_{\rm sph}$ correlation for the bulges of disk galaxies (S0+Sp).
However, using our dataset, 
we fit the $L_{\rm sph} - n_{\rm sph}$ relation for elliptical, lenticular and spiral galaxies separately, 
and found that the values of the slope and intercept for the lenticular galaxies 
are not consistent within the errors with those for the spiral galaxies, 
but are consistent within the errors with those for the elliptical galaxies.  
Given this, we conclude that in the $L_{\rm sph} - n_{\rm sph}$ diagram 
elliptical and lenticular galaxies form together a single (\emph{early-type}) sequence, 
whereas the combination of lenticular and spiral galaxies do not. 
According to the modified FITEXY routine, 
early-type galaxies follow\footnote{Using the BCES estimator, 
\cite{savorgnan2013} re-analyzed the dataset from \cite{grahamguzman2003} 
and obtained $L_{\rm sph} \propto n_{\rm sph}^{3.60 \pm 0.19}$ for the elliptical galaxies only. 
This result is in excellent agreement with the BCES linear regression obtained here for the early-type galaxies 
($L_{\rm sph} \propto n_{\rm sph}^{3.89 \pm 0.42}$) 
and, remarkably, it is exactly the same proportionality obtained here for the early-type galaxies with the modified FITEXY routine. } 
$L_{\rm sph} \propto n_{\rm sph}^{3.60 \pm 0.19}$, 
whereas late-type galaxies follow a shallower $L_{\rm sph} \propto n_{\rm sph}^{1.44 \pm 0.52}$ sequence.

@@

We have seen that the early- and late-type galaxies in the $M_{\rm BH} - n_{\rm sph}$ diagram can be fit with a single power-law, 
whereas they follow two separate correlations in the $M_{\rm BH} - L_{\rm sph}$ diagram (\emph{Paper II}). 
We now want to compare the amount of scatter among these two plots. 
In Table \ref{tab:intsc} we summarize the values of the intrinsic scatter 
in the $M_{\rm BH} - n_{\rm sph}$ and $M_{\rm BH} - L_{\rm sph}$ diagrams 
for all, early- and late-type galaxies, 
obtained with the modified FITEXY routine and the Bayesian estimator {\tt linmix\_err}. 
When considering all galaxies, irrespective of their morphological type, 
the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation is smaller than that of the $M_{\rm BH} - L_{\rm sph}$ relation. 
However, this is obviously not a fair comparison, 
because of the different nature of the $M_{\rm BH} - n_{\rm sph}$ and $M_{\rm BH} - L_{\rm sph}$ correlations 
(single and double power-law, respectively). 
One can obtain more informative results by considering early- and late-type galaxies separately. 
For the early-type galaxies, the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation is consistent\footnote{Looking 
at the results obtained with the modified FITEXY routine, 
the values of the intrisic scatter are only marginally consistent with each other, 
but looking at the results obtained with the Bayesian estimator {\tt linmix\_err}, 
the values of the intrisic scatter are fully consistent with each other. } 
with that of the $M_{\rm BH} - L_{\rm sph}$ relation (within their $1\sigma$ uncertainties). 
For the late-type galaxies, the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation is consistent with 
that of the $M_{\rm BH} - L_{\rm sph}$ relation, 
except for the inverse ($X|Y$) regression obtained with the modified FITEXY routine. \\
In passing, we note that the values of the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation 
are systematically smaller -- although consistent within the errors -- 
than the corresponding values of the intrinsic scatter of the $M_{\rm BH} - L_{\rm sph}$ relation. 
In addition, the values of the intrinsic scatter returned by the modified FITEXY routine are systematically smaller than 
those output by the Bayesian estimator {\tt linmix\_err}.

>>>>

\section{Data}
We populated the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams 
with the same galaxy sample used in \emph{Paper II} (and presented here in Table \ref{tab:sample}), 
i.e.~66 galaxies for which a dynamical measurement of the black hole mass has been reported in the literature 
(by \citealt{grahamscott2013} or \citealt{rusli2013}) 
and for which we were able to successfully model the light distribution and measure the spheroid structural parameters 
using $3.6\rm~\mu m$ \emph{Spitzer} satellite images. 
Our galaxy decompositions take into account bulge, disks, spiral arms, bars, rings, halo, 
extended or unresolved nuclear source and partially depleted core, 
and -- for the first time -- they were checked to be consistent with the galaxy kinematics 
\citep{atlas3dIII,scott2014,arnold2014}. 
Kinematical information was used to confirm the presence of disk components 
in the majority of early-type (elliptical + lenticular) galaxies and, more importantly,  
to estabilish the radial extent of these disks, 
which in most cases is not obvious from a visual inspection of the galaxy images. 
This enabled us to distinguish between intermediate-scale disks, 
that are fully embedded in the spheroid,  
and large-scale disks, that encase the bulge and dominate the light at large radii.  
\cite{ellicular} demonstrate that when an intermediate-scale disk is misclassified and modeled as a large-scale disk, 
the luminosity of the spheroid is underestimated, 
hence the galaxy incorrectly appears as a positive outlier (an ``over-massive'' black hole) in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
A detailed description of the dataset used here, the data reduction process and the galaxy modelling technique that we developed 
can be found in \emph{Paper I}, 
along with a discussion of how we estimated the uncertainties on the spheroid S\'ersic indices\footnote{The uncertainties associated with 
the spheroid S\'ersic indices were estimated with a method that takes into account systematic errors. 
This method consists in comparing, for each of our galaxies, the measurements of the spheroid S\'ersic index obtained by other studies 
with that obtained by us. 
Systematic errors are typically not considered by popular 2D fitting codes, which report only the statistical errors 
associated with their fitted parameters. 
Readers should refer to \emph{Paper I} for a more detailed discussion on this topic. }. 
For this analysis, we elected to use the spheroid S\'ersic indices obtained from the decomposition of the one-dimensional 
major-axis surface brightness profiles of our galaxies.  
The morphological classification (E = elliptical; E/S0 = elliptical/lenticular; S0 = lenticular; 
S0/Sp = lenticular/spiral; Sp = spiral; and ``merger'') follows from the galaxy decompositions illustrated in \emph{Paper I}. 
As in \emph{Paper II}, we will refer to early-type galaxies (E+S0) and late-type galaxies (Sp). 
The early-type bin includes the two galaxies classified as E/S0, 
whereas the two galaxies classified as S0/Sp and the two galaxies classified as ``mergers'' are included in neither the early- nor the late-type bin.

@@

\section{Analysis and results}
\label{sec:anal}
As in \emph{Paper II}, a linear regression analysis of the $L_{\rm sph} - n_{\rm sph}$ 
(Table \ref{tab:lregLn} and Figure \ref{fig:magn})
and $M_{\rm BH} - n_{\rm sph}$ (Table \ref{tab:lregMn} and Figure \ref{fig:mbhn}) diagrams 
was performed using three different routines: 
the BCES code from \cite{akritasbershady1996}, 
the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}.
All of these three routines take into account the intrinsic scatter, 
but only the FITEXY and the {\tt linmix\_err} codes allow one to quantify it.
\cite{tremaine2002} cautioned that 
the BCES estimator becomes ineffective when the dataset contains at least one low-precision measurement -- 
regardless of how many high-precision measurements are in the sample -- 
and tends to be biased in case of low number statistics 
or if the mean square of the uncertainties associated to the independent variable is comparable to 
the variance of the distribution of the independent variable. 
According to the results from the Monte Carlo Markov Chain simulations of \cite{tremaine2002} and \cite{novak2006}, 
these problems can be overcome with the use of the modified FITEXY routine. 
\cite{park2012} also concluded that the modified FITEXY routine performs better and returns less biased results than the BCES estimator, 
and noted that the the modified FITEXY routine is computationally less intensive than the Bayesian technique {\tt linmix\_err}. 
Given that at least one of our subsamples (the lenticular galaxies) has a small size 
and that the uncertainties associated to $n_{\rm sph}$ are relatively large 
compared to the range spanned by the $n_{\rm sph}$ values for most of our subsamples, 
we put more trust in the results obtained with the modified FITEXY routine 
and throughout the text we quote only those. \\ 
We report both symmetrical and nonsymmetrical linear regressions.  
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas nonsymmetrical forward ($Y|X$) regressions -- which minimize the scatter in the $Y$ direction -- 
allow one to predict the value of the observable $Y$ with the best possible precision. \\
We searched for extreme outliers in both the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams, 
and found that in our $L_{\rm sph} - n_{\rm sph}$ plot there are no $3\sigma$ outliers, 
whereas in our $M_{\rm BH} - n_{\rm sph}$ plot 
the lenticular galaxies NGC 0524 and NGC 3998 reside more than $3\sigma$ from the bisector linear regression for all galaxies. 
These two galaxies have therefore been excluded from the rest of the analysis. \\

\subsection{$L_{\rm sph} - n_{\rm sph}$}
Following \cite{graham2001bulges}, 
who showed that the $L_{\rm sph} - n_{\rm sph}$ relation is different for elliptical galaxies and the bulges of disk galaxies (S0+Sp), 
\cite{savorgnan2013} re-analyzed the data from \cite{grahamguzman2003} and \cite{grahamworley2008} 
and obtained two separate $L_{\rm sph} - n_{\rm sph}$ linear regressions for elliptical galaxies and the bulges of disk galaxies 
(in the B- and K-band, respectively). 
At the time, the $L_{\rm sph} - n_{\rm sph}$ datasets from \cite{grahamguzman2003} and \cite{grahamworley2008} were of the best quality available 
to investigate the $L_{\rm sph} - n_{\rm sph}$ relation for different galaxy morphological types. 
However, these datasets were not obtained from a homogeneous analysis, 
but they were a collection of results taken from various past bulge/disk decomposition studies. 
%It is also worth mentioning that, 
%although the K-band luminosities of the bulges of disk galaxies had been corrected for internal dust absorption, 
%{\bf such corrections are still not very accurate today?!? Driver et al. xxxx??? }  
Here we re-investigate the $L_{\rm sph} - n_{\rm sph}$ diagram using only our high-quality dataset. 
Spheroid luminositites and S\'ersic indices were obtained from accurate multicomponent decompositions,  
performed in a consistent manner using the $3.6\rm~\mu m$ band, which is less affected by dust extinction than the K-band. \\
\cite{grahamworley2008} presented a single $L_{\rm sph} - n_{\rm sph}$ correlation for the bulges of disk galaxies (S0+Sp).
However, using our dataset, 
we fit the $L_{\rm sph} - n_{\rm sph}$ relation for elliptical, lenticular and spiral galaxies separately, 
and found that the values of the slope and intercept for the lenticular galaxies 
are not consistent within the errors with those for the spiral galaxies, 
but are consistent within the errors with those for the elliptical galaxies.  
Given this, we conclude that in the $L_{\rm sph} - n_{\rm sph}$ diagram 
elliptical and lenticular galaxies form together a single (\emph{early-type}) sequence, 
whereas the combination of lenticular and spiral galaxies do not. 
According to the modified FITEXY routine, 
early-type galaxies\footnote{Using the BCES estimator, 
\cite{savorgnan2013} re-analyzed the dataset from \cite{grahamguzman2003} 
and obtained $L_{\rm sph} \propto n_{\rm sph}^{3.60 \pm 0.19}$ for the elliptical galaxies only. 
This result is in excellent agreement with the BCES linear regression obtained here for the early-type galaxies 
($L_{\rm sph} \propto n_{\rm sph}^{3.89 \pm 0.42}$) 
and, remarkably, it is exactly the same proportionality obtained here for the early-type galaxies with the modified FITEXY routine. } 
follow $L_{\rm sph} \propto n_{\rm sph}^{3.60 \pm 0.19}$, 
whereas late-type galaxies follow a shallower $L_{\rm sph} \propto n_{\rm sph}^{1.44 \pm 0.52}$ sequence.

@@

We have seen that the early- and late-type galaxies in the $M_{\rm BH} - n_{\rm sph}$ diagram 
can be fit (together) with a single power-law, 
whereas they follow two different correlations in the $M_{\rm BH} - L_{\rm sph}$ diagram (\emph{Paper II}). 
We now want to compare the amount of intrinsic scatter in these two plots. 
In Table \ref{tab:intsc} we report the values of the intrinsic scatter 
in the $M_{\rm BH} - n_{\rm sph}$ and $M_{\rm BH} - L_{\rm sph}$ diagrams 
for all, early- and late-type galaxies, 
obtained with the modified FITEXY routine and the Bayesian estimator {\tt linmix\_err}. 
When considering all galaxies, irrespective of their morphological type, 
the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation is smaller than that of the $M_{\rm BH} - L_{\rm sph}$ relation. 
However, this is obviously not a fair comparison, 
because of the different nature of the $M_{\rm BH} - n_{\rm sph}$ and $M_{\rm BH} - L_{\rm sph}$ correlations 
(single and double power-law, respectively). 
One can obtain more informative results by considering early- and late-type galaxies separately. 
For the early-type galaxies, the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation is consistent\footnote{Looking 
at the results obtained with the modified FITEXY routine, 
the values of the intrisic scatter are only marginally consistent with each other, 
but looking at the results obtained with the Bayesian estimator {\tt linmix\_err}, 
the values of the intrisic scatter are fully consistent with each other. } 
with that of the $M_{\rm BH} - L_{\rm sph}$ relation (within their $1\sigma$ uncertainties). 
For the late-type galaxies, the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation is consistent with 
that of the $M_{\rm BH} - L_{\rm sph}$ relation, 
except for the inverse ($X|Y$) regression obtained with the modified FITEXY routine. \\
In passing, we note that the values of the intrinsic scatter of the $M_{\rm BH} - n_{\rm sph}$ relation 
are systematically smaller -- although consistent within the errors -- 
than the corresponding values of the intrinsic scatter of the $M_{\rm BH} - L_{\rm sph}$ relation.  
In addition, the values of the intrinsic scatter returned by the modified FITEXY routine are systematically smaller than 
those output by the Bayesian estimator {\tt linmix\_err}.
